---
title: "Survival Differences by Tumor Stage in Breast Cancer: A Cox Proportional Hazards Analysis Using Real‑World Data"
subtitle: "This is a Report Template Quarto"
author: "Victoria Nguyen & Julaxis Love"
date: '`r Sys.Date()`'
format:
  html: 
    code-fold: true
course: Capstone Projects in Data Science
bibliography: references.bib # file contains bibtex for references
#always_allow_html: true # this allows to get PDF with HTML features
self-contained: true
execute: 
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

Slides: [slides.html](slides.html){target="_blank"} ( Go to `slides.qmd`
to edit)

## Introduction

Breast cancer is a major global health challenge, affecting millions of women and remaining one of the leading causes of cancer-related death. Although advances in screening, diagnosis, and treatment have improved survival for many patients, outcomes continue to vary widely. This variation is largely due to the heterogeneous nature of breast cancer, which differs across molecular subtypes, risk factors, and disease progression patterns [@Lukasiewicz2021BreastCancer; @Xiong2025BreastCancer]. As a result, early detection and personalized treatment strategies are critical for improving patient outcomes.

Recent literature highlights significant progress in understanding breast cancer biology, including the identification of molecular subtypes, prognostic biomarkers, and targeted therapies. Advances in molecular profiling and emerging tools such as AI-assisted diagnostics have improved the ability to tailor treatment to individual patients [@Lukasiewicz2021BreastCancer]. However, despite these innovations, tumor stage at diagnosis remains one of the most influential predictors of survival and continues to play a central role in clinical decision-making [@Xiong2025BreastCancer].

Understanding how tumor stage affects survival is therefore essential for both clinicians and patients. Survival analysis offers a robust framework for evaluating time-to-event outcomes while accounting for censored observations and varying follow-up times. Methods such as Kaplan–Meier survival curves and Cox proportional hazards models are widely used to examine survival patterns and assess the impact of clinical factors on patient prognosis.

Building on prior research, this study addresses the question: Does tumor stage at diagnosis influence survival time in breast cancer patients? We hypothesize that higher tumor stages are associated with shorter survival times. By examining this relationship, the study aims to reinforce the importance of early detection and accurate staging, while contributing evidence that supports personalized and stage-informed approaches to breast cancer care.


## Methods

-   Detail the models or algorithms used.

-   Justify your choices based on the problem and data.

*The common non-parametric regression model is*
$Y_i = m(X_i) + \varepsilon_i$*, where* $Y_i$ *can be defined as the sum
of the regression function value* $m(x)$ *for* $X_i$*. Here* $m(x)$ *is
unknown and* $\varepsilon_i$ *some errors. With the help of this
definition, we can create the estimation for local averaging i.e.*
$m(x)$ *can be estimated with the product of* $Y_i$ *average and* $X_i$
*is near to* $x$*. In other words, this means that we are discovering
the line through the data points with the help of surrounding data
points. The estimation formula is printed below [@R-base]:*

$$
M_n(x) = \sum_{i=1}^{n} W_n (X_i) Y_i  \tag{1}
$$$W_n(x)$ *is the sum of weights that belongs to all real numbers.
Weights are positive numbers and small if* $X_i$ *is far from* $x$*.*

*Another equation:*

$$
y_i = \beta_0 + \beta_1 X_1 +\varepsilon_i
$$

## Analysis and Results

### Data Exploration and Visualization

-   Describe your data sources and collection process.

-   Present initial findings and insights through visualizations.

-   Highlight unexpected patterns or anomalies.

A study was conducted to determine how...

```{r, warning=FALSE, echo=T, message=FALSE}
# loading packages 
library(tidyverse)
library(knitr)
library(ggthemes)
library(ggrepel)
library(dslabs)
```

```{python}
import pandas as pd
```

```{r, warning=FALSE, echo=TRUE}
# Load Data
kable(head(murders))

ggplot1 = murders %>% ggplot(mapping = aes(x=population/10^6, y=total)) 

  ggplot1 + geom_point(aes(col=region), size = 4) +
  geom_text_repel(aes(label=abb)) +
  scale_x_log10() +
  scale_y_log10() +
  geom_smooth(formula = "y~x", method=lm,se = F)+
  xlab("Populations in millions (log10 scale)") + 
  ylab("Total number of murders (log10 scale)") +
  ggtitle("US Gun Murders in 2010") +
  scale_color_discrete(name = "Region")+
      theme_bw()
  

```

### Modeling and Results

-   Explain your data preprocessing and cleaning steps.

-   Present your key findings in a clear and concise manner.

-   Use visuals to support your claims.

-   **Tell a story about what the data reveals.**

```{r}

```

### Conclusion

-   Summarize your key findings.

-   Discuss the implications of your results.

## References
