---
title: "Survival Differences by Tumor Stage in Breast Cancer: A Cox Proportional Hazards Analysis Using Real-World Data"
author: "Victoria Nguyen & Julaxis Love"
date: "`r Sys.Date()`"

format: 
  html:
    code-fold: true
    self-contained: true

course: Capstone Projects in Data Science
bibliography: references.bib

execute:
  warning: false
  message: false

editor:
  markdown:
    wrap: sentence
---

Slides: [slides.html](slides.html){target="_blank"}  
(Edit `slides.qmd` to modify presentation slides.)

---

## 1. Introduction

Breast cancer remains one of the most significant global health challenges and continues to rank among the leading causes of cancer-related mortality worldwide. Although advances in screening, diagnostic imaging, molecular profiling, and targeted therapies have substantially improved detection and treatment, survival outcomes still vary considerably across patients.

This variation reflects the heterogeneous nature of breast cancer, which differs in molecular subtype, biological aggressiveness, and patterns of disease progression. While innovations such as genomic assays and AI-assisted diagnostics have enhanced personalized treatment planning, traditional clinical indicators—particularly tumor stage at diagnosis—remain among the strongest predictors of survival.

Because tumor stage reflects the extent of disease spread at diagnosis, understanding its relationship with survival is critical for clinical decision-making and patient counseling.

This study investigates whether AJCC 6th Edition tumor stage is associated with overall survival time in breast cancer patients.  

**Hypothesis:** Higher tumor stages are associated with shorter survival times.

---

## 1.1 Overview of AJCC 6th Edition Breast Cancer Staging

The AJCC 6th edition staging system classifies breast cancer using:

- **T** – Tumor size  
- **N** – Regional lymph node involvement  
- **M** – Presence of distant metastasis  

These components are combined to assign an overall stage from I to IV.

### Stage I  
Localized tumors (≤ 2 cm) with minimal or no lymph node involvement.  
Typically associated with excellent survival outcomes.

### Stage II  
Larger tumors and/or limited regional lymph node involvement.  
Generally favorable survival but more variable than Stage I.

### Stage III  
Locally advanced disease involving multiple lymph nodes or adjacent structures.  
Significantly lower survival compared to Stages I and II.

### Stage IV  
Metastatic disease involving distant organs (e.g., bone, liver, lung, brain).  
Associated with substantially poorer survival outcomes.

Because stage at diagnosis strongly predicts prognosis, it serves as the primary predictor in this survival analysis.

---

## 1.2 Tumor Differentiation

Tumor differentiation describes how closely cancer cells resemble normal breast tissue under microscopic examination.

- **Well-differentiated**: Resemble normal cells; slower growth.
- **Moderately differentiated**: Intermediate characteristics.
- **Poorly differentiated**: Highly abnormal; more aggressive behavior.

Differentiation is closely related to tumor grade and provides additional biological context beyond anatomical staging. Two patients with identical AJCC stages may have different prognoses depending on tumor differentiation.

Including differentiation as a covariate allows for a more comprehensive survival analysis.

---

# 2. Methods

This project evaluates the relationship between AJCC 6th edition stage and overall survival using survival analysis techniques.

Survival analysis is appropriate when:
- The outcome is time-to-event.
- Some observations are censored.
- Follow-up time varies across individuals.

---

## 2.1 Survival Function and Kaplan–Meier Estimation

The survival function, \( S(t) \), quantifies the probability that an individual survives beyond a specified time \( t \). It provides a fundamental description of time-to-event outcomes in clinical studies.

The Kaplan–Meier estimator is used to estimate the survival function non-parametrically:

\[
\hat{S}(t) = \prod_{t_i \le t} \left(1 - \frac{d_i}{n_i}\right)
\]

Where:  
- \( t_i \) represents observed event times  
- \( d_i \) denotes the number of deaths at time \( t_i \)  
- \( n_i \) corresponds to the number of individuals at risk immediately prior to \( t_i \)  

Kaplan–Meier curves allow for visual comparison of survival distributions across categorical groups, such as tumor stage. Differences between curves are formally assessed using the **log-rank test**, which evaluates the null hypothesis of equivalent survival functions across groups.

---

## 2.2 Cox Proportional Hazards Model

To account for multiple covariates simultaneously, we applied the Cox proportional hazards model, a semi-parametric method widely used in clinical survival analysis:

\[
h(t \mid X) = h_0(t)\exp(\beta_1 X_1 + \beta_2 X_2 + \dots + \beta_p X_p)
\]

Where:  
- \( h_0(t) \) is the baseline hazard function representing the hazard when all covariates are zero  
- \( X \) denotes the vector of predictor variables (e.g., AJCC stage, age, tumor size, tumor differentiation)  
- \( e^{\beta} \) is the hazard ratio (HR) quantifying the effect of each covariate on the hazard  

Interpretation:  
- **HR > 1** indicates an increased risk of the event associated with the covariate  
- **HR < 1** indicates a decreased risk of the event  

This model allows adjustment for confounding factors and provides insight into the relative contribution of each clinical feature to overall survival.

---

## 2.3 Dataset and Analytical Workflow

Analyses were conducted using a real-world breast cancer dataset containing demographic, clinical, and tumor-specific variables. Key variables include:  
- **Outcome:** Survival time (months) and event status (death vs. censored)  
- **Primary predictor:** AJCC 6th edition tumor stage  
- **Covariates:** Age, race, tumor size, and tumor differentiation  

The analytical workflow comprised the following steps:  
1. Load and inspect the dataset for completeness and consistency.  
2. Prepare variables, including conversion of categorical variables to factors.  
3. Conduct exploratory data analysis to summarize patient demographics and tumor characteristics.  
4. Construct Kaplan–Meier survival curves stratified by AJCC stage and perform log-rank tests.  
5. Fit a Cox proportional hazards model to assess the association of tumor stage and covariates with survival, and evaluate proportional hazards assumptions using Schoenfeld residuals.  
6. Visualize results, including Kaplan–Meier curves and hazard ratio forest plots.

## 2.4 Analytical Steps

1. Create survival object using `Surv()`
2. Fit Kaplan–Meier curves using `survfit()`
3. Compare curves with `survdiff()`
4. Fit Cox model using `coxph()`
5. Extract hazard ratios and 95% CIs
6. Test proportional hazards using `cox.zph()`

---

## 2.5 Assumptions

- Proportional hazards assumption  
- Accurate survival measurement  
- Independent censoring  
- Correct model specification  

If violated, stratified or time-dependent models may be considered.

---

## 2.6 Software

All analyses were conducted in **R** using:

- `survival`
- `survminer`
- `dplyr`

# 3. Analysis and Results

```{r}
library(survival)
library(survminer)
library(dplyr)
library(ggplot2)
library(gt)
library(tidyr)
library(tidyverse)

data <- read.csv("breast_cancer_data.csv")

### cleaning data
cleaned_data <- data %>%
  select(
    Age, Race, Marital.Status, X6th.Stage, differentiate,
    Tumor.Size, Estrogen.Status, Progesterone.Status,
    Regional.Node.Examined, Reginol.Node.Positive, Survival.Months,
    Status
  )

cleaned_data <- cleaned_data %>%
  rename(
    `AJCC_Stage` = X6th.Stage,
    `Regional.Node.Positive` = Reginol.Node.Positive
  )

cleaned_data <- cleaned_data %>%
  mutate(
    Status = case_when(
      Status == "Alive" ~ 1,
      Status == "Dead"  ~ 0
    ),
    Race = case_when(
      Race == "White" ~ 1,
      Race == "Black" ~ 2,
      Race == "Other" ~ 3
    ),
    differentiate = case_when(
      differentiate == "Well differentiated" ~ 1, ##grade 1
      differentiate == "Moderately differentiated" ~ 2, ## grade 2
      differentiate == "Poorly differentiated" ~ 3, ## grade 3
      differentiate == "Undifferentiated" ~ 4
    ),
    Estrogen.Status = case_when(
      Estrogen.Status == "Positive" ~ 1,
      Estrogen.Status == "Negative" ~ 0
    ),
    Progesterone.Status = case_when(
      Progesterone.Status == "Positive" ~ 1,
      Progesterone.Status == "Negative" ~ 0
    ),
    Marital.Status = case_when(
      Marital.Status == "Married" ~ 1,
      Marital.Status == "Single " ~ 2,
      Marital.Status == "Divorced" ~ 3,
      Marital.Status == "Widowed" ~ 4,
      Marital.Status == "Separated" ~ 5
    )
  )

#------------------
variable_table <- data.frame(
  Variable = c(
    "Age", "Race", "6th Stage", "Differentiate", "Tumor Size",
    "Survival Months", "Estrogen Status", "Progesterone Status",
    "Regional Nodes Examined", "Regional Nodes Positive", "Status"
  ),
  Definition = c(
    "This variable is the patient’s age at diagnosis.",
    "This variable is the patient’s self-identified racial category.",
    "This variable is the cancer stage based on the AJCC 6th Edition.",
    "This variable is the tumor grade based on how abnormal the cells appear.",
    "This variable is the measured size of the primary tumor.",
    "This variable is the number of months from diagnosis to last follow-up or death.",
    "This variable is an indicator of estrogen receptor expression.",
    "This variable is an indicator of progesterone receptor expression.",
    "This variable is the number of lymph nodes examined.",
    "This variable is the number of lymph nodes found positive for cancer.",
    "This variable is the patient’s vital status at last follow-up."
  ),
  stringsAsFactors = FALSE
)

variable_table %>%
  gt() %>%
  tab_header(
    title = "Table 1. Variable Description"
  ) %>%
  tab_footnote(
    footnote = "Each variable in the dataset, accompanied by a qualitative description."
  )
```

**Explanation (Table 1):** This table provides a description of all variables in the dataset, helping viewers understand what each variable represents prior to analysis.

```
##Coding scheme
codebook <- tibble(
  Variable = c(
    "Status",
    "Race",
    "Differentiate(Tumor grade))",
    "Estrogen Status",
    "Progesterone Status",
    "Marital Status"
  ),
  
  Coding = c(
    "1 = Alive; 0 = Dead",
    "1 = White; 2 = Black; 3 = Other",
    "1 = Well differentiated; 2 = Moderately differentiated; 3 = Poorly differentiated; 4 = Undifferentiated",
    "1 = Positive; 0 = Negative",
    "1 = Positive; 0 = Negative",
    "1 = Married; 2 = Single; 3 = Divorced; 4 = Widowed; 5 = Separated"
  )
)

codebook %>%
  gt() %>%
  tab_header(
    title = "Table 2. Variable Coding Details"
  ) %>%
  cols_label(
    Variable = "Variable",
    Coding = "Coding Scheme"
  )
```
  **Explanation (Table 2):** This table shows the coding scheme used for categorical variables, which is critical for interpreting the results of survival analysis and Cox regression.

```
#Clean Data summary
make_cat_table <- function(data, var, label) {
  tbl <- table(data[[var]])
  pct <- prop.table(tbl) * 100
  
  tibble(
    Variable = label,
    Category = names(tbl),
    Summary = paste0(
      as.numeric(tbl), " (", sprintf("%.1f", pct), "%)"
    )
  )
}

# Build table sections
table_age <- tibble(
  Variable = "Age (mean ± SD)",
  Category = "",
  Summary = sprintf("%.1f ± %.1f",
                    mean(cleaned_data$Age, na.rm = TRUE),
                    sd(cleaned_data$Age, na.rm = TRUE))
)

table_race <- make_cat_table(cleaned_data, "Race", "Race")
table_marital <- make_cat_table(cleaned_data, "Marital.Status", "Marital Status")
table_stage <- make_cat_table(cleaned_data, "AJCC_Stage", "AJCC Stage")
table_grade <- make_cat_table(cleaned_data, "differentiate", "Tumor Grade")
table_er <- make_cat_table(cleaned_data, "Estrogen.Status", "Estrogen Status")
table_pr <- make_cat_table(cleaned_data, "Progesterone.Status", "Progesterone Status")
table_status <- make_cat_table(cleaned_data, "Status", "Vital Status")

table_tumor_size <- tibble(
  Variable = "Tumor Size (mean ± SD)",
  Category = "",
  Summary = sprintf("%.1f ± %.1f",
                    mean(cleaned_data$Tumor.Size, na.rm = TRUE),
                    sd(cleaned_data$Tumor.Size, na.rm = TRUE))
)

table_nodes_examined <- tibble(
  Variable = "Regional Nodes Examined (mean ± SD)",
  Category = "",
  Summary = sprintf("%.1f ± %.1f",
                    mean(cleaned_data$Regional.Node.Examined, na.rm = TRUE),
                    sd(cleaned_data$Regional.Node.Examined, na.rm = TRUE))
)

table_nodes_positive <- tibble(
  Variable = "Regional Nodes Positive (mean ± SD)",
  Category = "",
  Summary = sprintf("%.1f ± %.1f",
                    mean(cleaned_data$Regional.Node.Positive, na.rm = TRUE),
                    sd(cleaned_data$Regional.Node.Positive, na.rm = TRUE))
)

table_survival <- tibble(
  Variable = "Survival Months (mean ± SD)",
  Category = "",
  Summary = sprintf("%.1f ± %.1f",
                    mean(cleaned_data$Survival.Months, na.rm = TRUE),
                    sd(cleaned_data$Survival.Months, na.rm = TRUE))
)

# Combine all sections
table1 <- bind_rows(
  table_age,
  table_race,
  table_marital,
  table_stage,
  table_grade,
  table_tumor_size,
  table_er,
  table_pr,
  table_nodes_examined,
  table_nodes_positive,
  table_survival,
  table_status
)

# Create GT table
table1 %>%
  gt() %>%
  tab_header(
    title = "Table 3. Baseline Characteristics"
  ) %>%
  cols_label(
    Variable = "Variable",
    Category = "Category",
    Summary = "n (%) or Mean ± SD"
  ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = Variable
    )
  )
  
```
**Explanation (Table 3):** This table summarizes baseline characteristics of the cohort, including age, race, tumor stage, hormone status, and survival times. It provides an overview of patient demographics and clinical variables before analysis.

#### 3.2 Distributions/ Counts

#age
ggplot(cleaned_data, aes(Age)) +
  geom_histogram(binwidth = 5, fill = "darkgreen", color = "white") +
  labs(title = "Age Distribution", x = "Age", y = "Count")

  **Explanation (Figure 1):** This histogram shows the distribution of patient ages at diagnosis. Most patients cluster in middle age, which can influence survival outcomes.

#tumor
ggplot(cleaned_data, aes(Tumor.Size)) +
  geom_histogram(binwidth = 5, fill = "tomato", color = "white") +
  labs(title = "Tumor Size Distribution", x = "Tumor Size (mm)", y = "Count")
  
  **Explanation (Figure 2):** This histogram illustrates the distribution of primary tumor sizes. Larger tumor size at diagnosis may be associated with worse survival.

#hormone
ggplot(cleaned_data, aes(factor(Estrogen.Status))) +
  geom_bar(fill = "purple") +
  labs(title = "Estrogen Receptor Status", x = "Status (1=Positive, 0=Negative)", y = "Count")
  
  **Explanation (Figure 3):** This bar chart shows the number of patients with positive versus negative estrogen receptor status, an important predictor of treatment response and survival.

#stage
ggplot(cleaned_data, aes(AJCC_Stage)) +
  geom_bar(fill = "steelblue") +
  labs(title = "Figure 4.AJCC Stage Distribution", x = "AJCC Stage", y = "Count")
  
  **Explanation (Figure 4):** This figure displays the distribution of tumor stages at diagnosis. Stage is a key predictor of survival and is central to our analysis.

## 3.3 Kaplan–Meier Survival Analysis
Compare survival across AJCC stages<br>
Present Kaplan–Meier curves<br>
Report log-rank test results<br>

# -----------------------------
# Create survival object
# -----------------------------
surv_object <- Surv(cleaned_data$Survival.Months, cleaned_data$Status)

# Fit Kaplan–Meier model by AJCC stage
# -----------------------------
km_fit <- survfit(surv_object ~ AJCC_Stage, data = cleaned_data)

# -----------------------------
# Kaplan–Meier survival curves by stage
# -----------------------------
ggsurvplot(
  km_fit,
  data = cleaned_data,
  risk.table = TRUE,
  risk.table.height = 0.3,
  risk.table.fontsize = 5,
  pval = TRUE,
  conf.int = TRUE,
  legend.title = "AJCC Stage",
  xlab = "Months",
  ylab = "Survival Probability",
  palette = "Dark2",
  title = "Figure 5. Kaplan–Meier Survival by AJCC Stage"
)

**Explanation (Figure 5):** Kaplan–Meier curves compare survival across AJCC stages. The plot and log-rank test show whether survival differs significantly between stages.

## Kaplan–Meier survival curves by estrogen
km_fit_er <- survfit(surv_object ~ Estrogen.Status, data = cleaned_data)

ggsurvplot(
  km_fit_er,
  data = cleaned_data,
  risk.table = TRUE,
  pval = TRUE,
  conf.int = TRUE,
  legend.title = "ER Status",
  legend.labs = c("Negative", "Positive"),
  title = "Figure 6. Kaplan–Meier Survival by Estrogen Receptor Status"
)

**Explanation (Figure 6):** These survival curves show differences in overall survival based on estrogen receptor status. Positive ER status is generally associated with better prognosis.

## 3.4 Cox Proportional Hazards Model
Fit Cox model adjusting for covariates<br>
Report hazard ratios and confidence intervals<br>
Evaluate proportional hazards assumption<br>
```{r}
# -----------------------------
# Fit Cox proportional hazards model
# -----------------------------
cox_model <- coxph(
  Surv(Survival.Months, Status) ~ AJCC_Stage + Age + Race + differentiate + Tumor.Size,
  data = cleaned_data
)

# -----------------------------
# View Cox model summary
# -----------------------------
summary(cox_model)

# -----------------------------
# Forest plot of hazard ratios
# -----------------------------
ggforest(cox_model, data = cleaned_data, main = "Figure 7. Hazard Ratios for Breast Cancer Survival")
```

**Explanation (Figure 7):** This forest plot visualizes hazard ratios from the Cox proportional hazards model, adjusting for age, race, tumor grade, and tumor size. HR > 1 indicates increased risk of death. The proportional hazards assumption is tested to validate model reliability.
```
# -----------------------------
# Test proportional hazards assumption
# -----------------------------
ph_test <- cox.zph(cox_model)
print(ph_test)
plot(ph_test)
```
## 4. Discussion
- Interpretation of findings
- Comparison with existing literature
- Clinical implications
- Study limitations

---

# 5. Conclusion

Higher AJCC tumor stages are expected to demonstrate progressively worse survival outcomes. Confirming this relationship reinforces the importance of early detection and accurate staging in breast cancer management.

---

# References
Ali S, Hamam D, Liu X and Lebrun J-J (2022) Terminal differentiation and anti-tumorigenic effects of prolactin in breast cancer. Front. Endocrinol. 13:993570. doi: 10.3389/fendo.2022.993570

Koh, J., & Kim, M. J. (2019). Introduction of a New Staging System of Breast Cancer for Radiologists: An Emphasis on the Prognostic Stage. Korean journal of radiology, 20(1), 69–82. https://doi.org/10.3348/kjr.2018.0231

Nadia Howlader, Kathleen A. Cronin, Allison W. Kurian, Rebecca Andridge; Differences in Breast Cancer Survival by Molecular Subtypes in the United States. Cancer Epidemiol Biomarkers Prev 1 June 2018; 27 (6): 619–626. https://doi.org/10.1158/1055-9965.EPI-17-0627

Bewick, V., Cheek, L., & Ball, J. (2004). Statistics review 12: survival analysis. Critical care (London, England), 8(5), 389–394. https://doi.org/10.1186/cc2955

Bradburn, M. J., Clark, T. G., Love, S. B., & Altman, D. G. (2003). Survival analysis part II: multivariate data analysis--an introduction to concepts and methods. British journal of cancer, 89(3), 431–436. https://doi.org/10.1038/sj.bjc.6601119

Xu, M., Shan, D., Zhang, R., Li, J., Guo, L., Chen, X., & Qu, J. (2025). Differentiation of breast cancer subtypes and correlation with biological status using functional magnetic resonance imaging: comparison with amide proton transfer-weighted imaging and diffusion-weighted imaging. Quantitative imaging in medicine and surgery, 15(7), 6102–6117. https://doi.org/10.21037/qims-24-2174

Breast Cancer Stages | Understanding Breast Cancer Staging. (2025, December 2). Cancer.gov. https://www.cancer.gov/types/breast/stages ‌