---
title: "Survival Differences by Tumor Stage in Breast Cancer"
subtitle: "A Cox Proportional Hazards Analysis Using Real-World Data"
author: "Victoria Nguyen & Julaxis Love (Advisor: Dr. Cohen)"
date: '`r Sys.Date()`'
format:
  revealjs:
    theme: dark
    slide-number: true
    transition: fade
    css: custom.scss
course: Capstone Projects in Data Science
bibliography: references.bib # file contains bibtex for references
#always_allow_html: true # this allows to get PDF with HTML features
self-contained: true
execute: 
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
library(survival)
library(survminer)
library(dplyr)
library(ggplot2)

# Load data
data <- read.csv("breast_cancer_data.csv")

# Clean Data
cleaned_data <- data %>%
  select(
    Age, Race, Marital.Status, X6th.Stage, differentiate,
    Tumor.Size, Estrogen.Status, Progesterone.Status,
    Regional.Node.Examined, Reginol.Node.Positive,
    Survival.Months, Status
  ) %>%
  rename(
    AJCC_Stage = X6th.Stage,
    Regional.Node.Positive = Reginol.Node.Positive
  ) %>%
  mutate(
    Status = case_when(
      Status == "Alive" ~ 1,
      Status == "Dead" ~ 0
    ),
    differentiate = case_when(
      differentiate == "Well differentiated" ~ 1,
      differentiate == "Moderately differentiated" ~ 2,
      differentiate == "Poorly differentiated" ~ 3,
      differentiate == "Undifferentiated" ~ 4
    ),
    Race = case_when(
      Race == "White" ~ 1,
      Race == "Black" ~ 2,
      Race == "Other" ~ 3
    ),
    Estrogen.Status = case_when(
      Estrogen.Status == "Positive" ~ 1,
      Estrogen.Status == "Negative" ~ 0
    ),
    Progesterone.Status = case_when(
      Progesterone.Status == "Positive" ~ 1,
      Progesterone.Status == "Negative" ~ 0
    )
  )

# survival objects 
surv_object <- Surv(cleaned_data$Survival.Months, cleaned_data$Status)

km_fit <- survfit(surv_object ~ AJCC_Stage, data = cleaned_data)

km_fit_er <- survfit(surv_object ~ Estrogen.Status, data = cleaned_data)

cox_model <- coxph(
  Surv(Survival.Months, Status) ~ AJCC_Stage + Age + Race + differentiate + Tumor.Size,
  data = cleaned_data
)

ph_test <- cox.zph(cox_model)
```

## Introduction {.smaller}

- Breast cancer survival varies significantly across patients  
- Understanding what drives survival differences is critical  
- Tumor stage is a key clinical predictor  
- Outcomes are influenced by multiple factors  

**Goal:**  
Evaluate how tumor stage impacts survival while adjusting for clinical and demographic variables  

---

## Research Question

- Does tumor stage significantly impact survival?  
- How does risk change across stages?  
- Do other variables (age, tumor size, differentiation) matter?  

**Hypothesis:**  
Higher tumor stage → higher risk of death  

---

## Data Overview

- Real-world breast cancer dataset  
- Key variables:
  - Survival time (months)
  - Vital status (alive/dead)
  - AJCC stage (I–IV)
  - Age, race
  - Tumor size & differentiation
  - Hormone receptor status  

---

## Methods

We used survival analysis techniques:

- Kaplan–Meier curves → compare survival across groups  
- Log-rank test → test differences in survival  
- Cox proportional hazards model → estimate effect of covariates  

---

## Kaplan–Meier Survival by AJCC Stage

```{r echo=FALSE}
ggsurvplot(
  km_fit,
  data = cleaned_data,
  risk.table = TRUE,
  pval = TRUE,
  conf.int = TRUE,
  title = "Survival by AJCC Stage"
)
```
### Key Insights
- Survival decreases as stage increases
- Clear separation between stages

## Survival by Estrogen Status
```{r}
ggsurvplot(
  km_fit_er,
  data = cleaned_data,
  risk.table = TRUE,
  pval = TRUE,
  conf.int = TRUE,
  legend.labs = c("Negative", "Positive"),
  title = "Survival by Estrogen Receptor Status"
)
```
### Key Insights
- ER-positive patients show better survival
- Hormone status is an important prognostic factor

## Cox Proportional Hazards Model
### Adjusted for: 
- Tumor stage
- Age
- Race
- Tumor size
- Differentiation

## Cox Proportional Hazards Model continued
```{r echo=FALSE, fig.width=7, fig.height=4.5}
summary(cox_model)
ggforest(cox_model, data = cleaned_data)
```
## Cox Proportional Hazards Model continued
### Key Insights
- Tumor stage is the strongest predictor of survival
- Higher stage → increased hazard of death

## Proportional Hazards Assumption
```{r}
{echo=FALSE}
ph_test
plot(ph_test)
```
## Proportional Hazards Assumption continued
### Key Insights
- No major violations detected
- Model assumptions are satisfied
---
## Key Findings
- Higher tumor stage increases risk of death
- Tumor size and differentiation are significant predictors
- ER-positive status associated with improved survival
- Results remain significant after adjustment

## Discussion
- Tumor stage is a strong independent predictor of survival
- Survival is influenced by multiple clinical factors
- Findings align with clinical expectations

## Conclusion
- Tumor stage strongly impacts survival outcomes
- Multivariable modeling improves risk estimation
- Survival analysis provides meaningful clinical insights

## Implications
- Importance of early detection
- Supports personalized treatment planning
- Helps identify high-risk patients
